Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-06-13 manuscript.Rmd
revisions

Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate. First, the magnitude of movement can vary greatly *within* a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in "tummy time", or they can crawl in a myriad of ways [@AdolphVereijken1998]. Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a "V" position with legs fully extended, or in a "W" position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso. Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019]. Although @GreenspanCunha2021 were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground. Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [@Preece2009; @Arif2015], children [@NamPark2013; @RenDing2016; @StewartNarayanan2018], and infants [@YaoPlotz2019; @FranchakScott2021; @AiraksinenRasanen2020]. Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated. Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording. As described below, the video recording was not continuous. The camera created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).

@@

Because the camera only had the battery life to record for ~3 hours (split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was automatically recorded, termed the *distal comparison*. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of ~40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of ~5 s in either direction.

@@

We report three sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r round(sum(duration$duration_hours))` hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped. This leads to two final questions---how successfully did recordings capture infants' entire day, and how well do findings from full-day classification converge with findings that use other methodologies?

>>>>

Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate. First, the magnitude of movement can vary greatly *within* a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in "tummy time", or they can crawl in a myriad of ways [@AdolphVereijken1998]. Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a "V" position with legs fully extended, or in a "W" position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso. Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity [@KwonZavos2019; @PatelShi2019]. Although @GreenspanCunha2021 were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground. Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [@Preece2009; @Arif2015], children [@NamPark2013; @RenDing2016; @StewartNarayanan2018], and infants [@YaoPlotz2019; @FranchakScott2021; @AiraksinenRasanen2020]. Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated. Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotations.

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019]. Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest.

Video recordings were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary "Boosted Battery Base" attached to the action camera to allow for a total of 3 hours of recording. As described below, the video recording was not continuous. The camera created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).

@@

Because the camera only had the battery life to record for ~3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \@ref(fig:exemplar-timeline)), the *video period* comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the *proximal comparison*, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was automatically recorded, termed the *distal comparison*. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of ~40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of ~5 s in either direction.

@@

We report three sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r round(sum(duration$duration_hours))` hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides, to our knowledge, the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The algorithm was not designed to classify behavior during transportation (e.g., strollers, automobiles) so data collection was restricted to in-home behavior. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped. This leads to two final questions---how successfully did recordings capture infants' entire day, and how well do findings from full-day classification converge with findings that use other methodologies?
